---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fishglobr

<!-- badges: start -->
<!-- badges: end -->

## THIS IS AN EXPERIMENTAL DRAFT PROJECT NOT BE USED AT THIS STAGE!

fishglobr (perhaps to be renamed to fishglob) is an R package for efficiently downloading, caching, and querying the [FishGlob](https://fishglob.sites.ucsc.edu/) database. FISHGLOB is an international consortium of scientists, experts and data providers who collect, curate, standardize, share and analyse data from scientific bottom trawl surveys.

## Installation

You can install the development version of fishglobr like so:

``` r
# temporary location; to be moved to fishglob group
pak::pak("seananderson/fishglobr")
```

## Example

First we load the package and download and install the database locally.
We only have to do this once.
The data is stored in a local DuckDB database.

```{r load, message=FALSE, eval=FALSE}
library(fishglobr)
install_fishglob_data()
#> Downloading catch ■■■■■■■■■■■■■■■■■■■■■■■■■■■■■■■ |  65 MB/ 65 MB ETA:  0s
#> Downloading haul ■■■■■■■■■■■■■■■■■■■■■■■■■■■■■■■ | 4.7 MB/4.7 MB ETA:  0s
#> Downloading taxon ■■■■■■■■■■■■■■■■■■■■■■■■■■■■■■■ |  49 kB/ 49 kB ETA:  0s
#> Downloading survey ■■■■■■■■■■■■■■■■■■■■■■■■■■■■■■■ | 0.7 kB/0.7 kB ETA:  0s
#> Building the local FishGlob DuckDB database
#> FishGlob data are ready in
#> ~/Library/Caches/org.R-project.R/R/fishglobr/fishglob.duckdb
```

```{r, echo=FALSE}
library(fishglobr)
```

```{r cars}
# Find available agency data sources
get_sources()

# The full list of available species:
spp <- get_scientific_names()
spp[1:5]

# All Canary Rockfish data held by DFO
dat <- get_data(
  scientific_name = "sebastes pinniger",
  source = "DFO"
)
nrow(dat)

# All Sebastes data globally from 2000 to 2021
# using a regular expression
dat <- get_data(
  scientific_name = "^sebastes*",
  regex = TRUE,
  year = 2000:2021
)
head(dat[c("scientific_name", "year", "latitude", "longitude", "wgt_cpue")], n = 3)
```
